\begin{acknowledgements}

First I would like to thank my supervisors, Tim French and Rowan Davies, for their incredible support throughout my postgraduate studies and even earlier.
It was thanks to the encouragement of Tim and Rowan that I attended the Logic Summer School at the Australian National University in 2010, beginning my journey in formal logic.
Since then I have attended numerous academic conferences and events, and I have Tim and Rowan to thank for their advice and feedback in writing my academic publications to date, for their financial support in attending these events on other continents, and for their advice, support and encouragement more generally during my studies.

I would like to thank Hans van Ditmarsch and Sophie Pinchinat for collaborating with me, and for graciously inviting me to work with them for several months in France in 2014.
I would additionally like to thank Hans for his financial support during my two month visit at LORIA in 2014.
I would also like to thank the anonymous reviewers of M4M 2011, AiML 2012 and 2014, and LICS 2013, whose helpful recommendations found their way into my academic publications and ultimately into my thesis.

I would like to acknowledge the financial support of the Prescott Postgraduate Scholarship, that has supported all of my postgraduate studies, the Graduate Research School Student Travel Award, that supported my attendance of ESSLLI and AiML in 2012, and the Convocation Postgraduate Research Travel Award, that supported my attendance of AiML in 2014.

Finally I would like to thank my significant other, Clara, for being my source of emotional stability during a stressful period of my life, completing my thesis.

\end{acknowledgements}
