\chapter{Refinement modal logic}\label{rml}

In this chapter we recall the refinement modal logic (\logicRml{}) of van Ditmarsch and French~\cite{vanditmarsch:2009} and provide semantic results about refinements and refinement modal logic.
\logicRml{} is an extension of modal logic that quantifies over the refinements of a Kripke model.
In epistemic settings refinements correspond to the results of a very general notion of epistemic updates.
However unlike public announcements or action models, refinements in general are not backed by a model or operation for epistemic updates that produces the results.  
In contrast to previous treatments of \logicRml{}~\cite{vanditmarsch:2009,vanditmarsch:2010}, which considered \logicRml{} specifically in the setting of \classK{}, our treatment considers \logicRml{} in different settings, including multi-agent doxastic and epistemic logics.
In the present chapter we provide definitions and results common to all or most of the settings that we consider.
In Section~\ref{rml-refinements} we recall the definition of refinements, we provide a partial characterisation of refinements through the preservation of the validity of positive formulas, we show a correspondence between the refinements of finite Kripke models and the results of executing action models, and we show many related results.
In Section~\ref{rml-semantics} we recall the semantics of \logicRml{} and provide semantic results common to the settings that we consider.
In following chapters we consider \logicRml{} in the settings of \classK{}, \classKFF{}, \classKD{}, \classS{}, and \classKF{}, providing results specific to each setting.

\input{rml-refinements.tex}
\input{rml-semantics.tex}
